\documentclass[a4paper,10pt]{report}
\input{macros_docu} % Customizations 

% ---- format de page A4
	\setlength{\textwidth }{16cm}	% largeur de ligne
	\setlength{\textheight}{23cm}   % hauteur du texte
	\setlength{\oddsidemargin}{0cm} % marge pages impaires
	\setlength{\evensidemargin}{0cm}% marge pages paires
	\setlength{\topmargin}{0cm} 	

% Title Page
\title{Sofa Documentation}
\author{The Sofa Team}



\begin{document}
\maketitle

\begin{abstract}
This is the Sofa documentation
\end{abstract}

\tableofcontents

% This chapter is borrowed from a stand-alone document defined in directory introduction/
\chapter{Introduction to \sofa}
\graphicspath{{introduction/}}  % to include images
\input{introduction/introduction_body}

% This chapter is borrowed from a stand-alone document defined in directory design/
\chapter{Design}
\graphicspath{{design/}}  % to include images
\input{design/design_body}

% This chapter is defined here
\chapter{How to contribute to this documentation}
\section{Document structure}
This document gathers the content of other documents located in different subdirectories. These documents can also be compiled as stand-alone documents.
The structure can be illustrated as follows:
\begin{itemize}
 \item \texttt{sofadocumentation.tex} : the root document.
 \item \texttt{macros\_docu.tex} : custom commands and macros. This file is included by the root document.
 \item \texttt{introduction} : a subdirectory containing a chapter of this document.
\begin{itemize}
 \item \texttt{introduction.tex} : a stand-alone article containing the same. This file may include \texttt{../macros\_docu.tex} to use the common custom commands.
 \item \texttt{introduction\_body.tex} : the text of the chapter/article. This file is included as a chapter by \texttt{sofadocumentation.tex}, and included as full article text by \texttt{introduction.tex}.
\end{itemize}
 \item there could (and hopefully will !) be other subdirectories with a similar structure.
\end{itemize}

\section{Compiling the document}
\subsection{File formats}
The graphics are handled by: \texttt{\textbackslash usepackage[pdftex]\{graphicx\}}. This allows the inclusion of \texttt{.png} images rather than \texttt{.eps}, which makes the image files much smaller, and compilation of \texttt{html} faster. The result of the compilation is a \texttt{.pdf} rather than a \texttt{.dvi} file.

\subsection{Include paths}
The root document includes files in the subdirectories, which in turn include files too. The problem is that the path from the subdirectory (used when compiling a stand-alone article) is not the same as from the parent directory (used when compiling the whole report).
It seems that \LaTeX has no include path command to circumvent this problem.

Fortunately, \LaTeX uses an environment variable named TEXINPUTS which represents a list of directories to search. We use this variable to allow file inclusion, by adding the parent directories to the list. The command lines depend on the shell, e.g.:
\begin{itemize}
 \item using bash: \texttt{export TEXINPUTS=\$\{TEXINPUTS\}:..:../.. ; latex sofadocumentation.tex}
 \item using tcsh: \texttt{setenv TEXINPUTS \$\{TEXINPUTS\}:..:../.. ; latex sofadocumentation.tex}
\end{itemize}

Using the \texttt{Kile} graphical front-end, you can customize the command line in menu \texttt{Settings/Kile/Tools/Build}, select tool \texttt{LATEX} and write the appropriate command line in the \texttt{Command} field.

\subsection{HTML}
HTML can be generated using the following command:\\
\texttt{latex2html sofadocumentation.tex -mkdir -dir ./html -show\_section\_numbers -split 1}
\\
Currently, the listings do not appear in html.

\end{document}          
